Binary Latent Representations for Efficient Ranking

### 2017-06-16 paper/paper.tex

We use two loss functions to train the model:
\begin{itemize}
\item Bayesian personalised ranking (BPR, \citet{rendle2009bpr}), and
\item adaptive sampling maximum margin loss, following \citet{weston2011wsabie}.
\end{itemize}
For both loss functions, for any known positive user-item interaction pair $(u, i)$, we sample an implicit negative item $j \in \symUserInteractionSet^-$. For BPR, the loss for any such triplet is given by
\begin{equation}
1 - \sigma\left(r_{ui} - r_{uj}\right),
\end{equation}
where $\sigma$ denotes the sigmoid function.
The adaptive sampling loss is given by
\begin{equation}
\abs{1 - r_{ui} + r_{uj}}_{+}.
\end{equation}
For any $(u, i)$ pair, if the sampled negative item $j$ results in a zero loss (that is, the desired pairwise ordering is not violated), a new negative item is sampled, up to a total of $k$ attempts. This leads the model to perform more gradient updates in areas where its ranking performance is poorest.

@@

Results on scoring speed and memory use broadly confirm the naive calculations from section \ref{sec:approach}, converging to over 20 times faster scoring at around 3\% of memory as latent dimensionality increases.
As expected, for the same dimensionality a binary model achieves lower ranking accuracy. On average, the accuracy loss is around 8\%, varying between 2\% and 16\%.
Nevertheless, the model's speed enables some very favourable tradeoffs.

>>>>

We use two loss functions to train the model:
\begin{itemize}
\item Bayesian personalised ranking (BPR, \citet{rendle2009bpr}), and
\item adaptive sampling maximum margin loss, following \citet{weston2011wsabie}.
\end{itemize}
For both loss functions, for any known positive user-item interaction pair $(u, i)$, we uniformly sample an implicit negative item $j$ \in \symUserInteractionSet^-$. For BPR, the loss for any such triplet is given by
\begin{equation}
1 - \sigma\left(r_{ui} - r_{uj}\right),
\end{equation}
where $\sigma$ denotes the sigmoid function.
The adaptive sampling loss is given by
\begin{equation}
\abs{1 - r_{ui} + r_{uj}}_{+}.
\end{equation}
For any $(u, i)$ pair, if the sampled negative item $j$ results in a zero loss (that is, the desired pairwise ordering is not violated), a new negative item is sampled, up to a total of $k$ attempts. This leads the model to perform more gradient updates in areas where its ranking performance is poorest.

@@

Results on scoring speed and memory use broadly confirm the naive calculations from section \ref{sec:approach}, converging to over 20 times faster scoring at around 3\% of memory as latent dimensionality increases. Memory savings plateau as latent dimensionality increases and storing the binary scaling factors becomes less important.

As expected, for the same dimensionality a binary model achieves lower ranking accuracy. On average, the accuracy loss is around 8\%, varying between 2\% and 16\%.
Nevertheless, the model's speed enables some very favourable tradeoffs.
